Stage-Splitting a Modal Language

### 2014-06-26 paper/semantics.tex
changed defn of hold

To understand how this behavior can be achieved, consider evaluation of the stage-\bbtwo\ term below, which contains nested $\next$ and $\prev$ blocks:
\begin{lstlisting} 
prev{
  let x = (next {1+2}, 3+4) in
  next{ prev{#1 x} * prev{#1 x} * prev{hold(#2 x)} }
}
\end{lstlisting}
Our goal is to evaluate away all of the stage-\bbone\ portions of this code, but there is more than one reasonable way to do this.
One option, taken by \cite{davies96}, is to treat \verb|next{1+2}| as a value for the purposes of stage-\bbone~evaluation, 
consequently duplicating it during substitution for {\tt x}.  
This produces:
\begin{lstlisting}
(1+2) * (1+2) * 7
\end{lstlisting}

@@

Achieving this behavior mechanically requires us to resolve a contradiction:
we must substitute for \texttt{x} at stage \bbone, but we cannot evaluate inside the $\next$ block bound to it. 
Our solution is to replace the $\next$ with a new variable and create an explicit substitution (shown with a $\mapsto$) binding that variable to the $\next$'s contents.  
This substitution then floats up to the top of the containing $\prev$ block:
\begin{lstlisting} 
prev {
[y|->1+2]
  let x = (yhat, 7) in
  next{prev{#1 x} * prev{#1 x} * prev{hold (#2 x)}}
}
\end{lstlisting}
Note that the variable left behind is rendered with a hat.  By treating this {\em hatted variable} as a value, we're now free to perform the stage-\bbone~substitution for {\tt x}.  
\begin{lstlisting} 
prev {
[y|->1+2]
  next{
    prev{#1 (yhat, 7)} * prev{#1 (yhat, 7)} *
    prev{hold (#2 (yhat, 7))}
  }
}
\end{lstlisting}
To evaluate the remaining $\next$, we must first partially evaluate the body by finding all of the contained stage-\bbone~terms and reducing them. 
As a rule, these will reduce to hatted variables, and $\prev$ removes the hat, leaving the variable in place:
\begin{lstlisting} 
prev {
[y|->1+2]
    next{ y * y * 7 }
}
\end{lstlisting}
Once again, we lift the contents of the $\next$ into a substitution:
\begin{lstlisting} 
prev {
[y|->1+2]
[z|->y*y*7]
    zhat
}
\end{lstlisting}
Finally, when evaluating the outer $\prev$, we must {\em reify} the contained substitutions into let statements, yielding
\begin{lstlisting} 
let y = 1+2 in
let z = y * y * 7 in z
\end{lstlisting}

@@

%	prev{ hold (2+4) }

>>>>

To understand how this behavior can be achieved, consider evaluation of the stage-\bbtwo\ term below, which contains nested $\next$ and $\prev$ blocks:
\begin{lstlisting} 
prev{
  let x = (next {1+2}, 3+4) in
  next{ prev{#1 x} * prev{#1 x} * hold{#2 x} }
}
\end{lstlisting}
Our goal is to evaluate away all of the stage-\bbone\ portions of this code, but there is more than one reasonable way to do this.
One option, taken by \cite{davies96}, is to treat \verb|next{1+2}| as a value for the purposes of stage-\bbone~evaluation, 
consequently duplicating it during substitution for {\tt x}.  
This produces:
\begin{lstlisting}
(1+2) * (1+2) * 7
\end{lstlisting}

@@

Achieving this behavior mechanically requires us to resolve a contradiction:
we must substitute for \texttt{x} at stage \bbone, but we cannot evaluate inside the $\next$ block bound to it. 
Our solution is to replace the $\next$ with a new variable and create an explicit substitution (shown with a $\mapsto$) binding that variable to the $\next$'s contents.  
This substitution then floats up to the top of the containing $\prev$ block:
\begin{lstlisting} 
prev {
[y|->1+2]
  let x = (yhat, 7) in
  next{prev{#1 x} * prev{#1 x} *  hold{#2 x}}
}
\end{lstlisting}
Note that the variable left behind is rendered with a hat.  By treating this {\em hatted variable} as a value, we're now free to perform the stage-\bbone~substitution for {\tt x}.  
\begin{lstlisting} 
prev {
[y|->1+2]
  next{
    prev{#1 (yhat, 7)} * prev{#1 (yhat, 7)} *
    hold{#2 (yhat, 7)}
  }
}
\end{lstlisting}
To evaluate the remaining $\next$, we must first partially evaluate the body by finding all of the contained stage-\bbone~terms and reducing them. 
As a rule, these will reduce to hatted variables, and $\prev$ removes the hat, leaving the variable in place:
\begin{lstlisting} 
prev {
[y|->1+2]
    next{ y * y * 7 }
}
\end{lstlisting}
Once again, we lift the contents of the $\next$ into a substitution:
\begin{lstlisting} 
prev {
[y|->1+2]
[z|->y*y*7]
    zhat
}
\end{lstlisting}
Finally, when evaluating the outer $\prev$, we must {\em reify} the contained substitutions into let statements, yielding
\begin{lstlisting} 
let y = 1+2 in
let z = y * y * 7 in z
\end{lstlisting}

@@

%	hold{2+4}
